A flexible framework to prototype, develop and manage ubiquitous applications for intelligent environments

### 2013-10-02 trunk/ labtempo-ubicomp-projects/UbicompMiddleware/doc/PerCom2014/percom2014_conf.tex
more about communication including two new figures change position of subsection Structure from Resource Agent section trying to write algorithm to SmartLiC changed position of first figure

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

@@

% With the purpose of evaluating the proposal we implemented the concepts 
% and enable the context interpretation and 
%  even without the need of purchasing all those devices
%  (with a touch screen interface)
% The abstractions provided, and the management services conceived, allow developers to handle resources spread within the intelligent environment in a simple and homogeneous way, and enhances the capability of implementing context sensible applications, while still keep the flexibility on developing such applications. 
% Above the platform we implemented a management and prototyping interface that uses a tablet device to allow the testing of applications running on different configurations of the environment, which may be loaded with simulators or the representation of the resources. The interface also allows the composition of context rules either by developers or non trained users.
%FIXME: ``while still keep the flexibility ON developing such applications'' <- este on está certo?

@@

% [architecture picture]
\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{figures/architecture-refactored-v2.pdf}
\caption{Framework Layer Architecture}
\label{fig:arch}
\end{figure}

@@

@@

%[David]
The conceptual framework proposed at this work was tested through a reference implementation called \mbox{SmartAndroid} Project\footnote{For more information visit www.tempo.uff.br/smartandroid}. The project includes a programming API to help the developers in constructing UbiApps as well as offers basic management services to provide resource registration, discovery and localization. The SmartAndroid has been developed over Google Android platform as we believe that in a near future it will be widely used in sensors and devices embedded in furnitures and home appliances.

[we have implemented most of the framework features, thought much more remain to be done]

@@

  \subsection{Context Rules Parsing} \label{subsec:rules}
  [implementation of context rules]
  %   Support the interpretation of the context framework enables a separation of concerns, where developers abstract implementation of contextual rules and focus on the application logic.

  \subsection{Context Rules Running}

@@

  \subsection{Interface for Management and Prototyping} \label{subsec:ipgap}
  % interface de prototipagem e gerenciamento de aplicações pervasivas
  % prototyping and management of pervasive applications interface
  % pervasive applications' management and prototyping interface
  % PAMPI
  % management and prototyping interface MPI
  [description]\\
  %[David]
The main purpose of the [IPGAP] is provide to developers a hybrid environment where simulated and real resources can interact to each other transparently, enabling that their UbiApps are tested and evaluated through a wide variety of scenarios. RAs from sensors, actuators and household appliances can be graphically represented at our tool, along with a schematic map from the intelligent environment so that the developer can visualize the context of all resources and debug his application.

Besides, non-technical users are able to use [IPGAP] to manage their intelligent environment remotely using a mobile device such as a tablet. Through a intuitive GUI, he can not only visualize the state of their home devices but control them, for instance, turning off a stove burner. By touching the device in the screen, an interface that presents its main functions is shown, where the user can operate it as the tool were a remote control.

  [main figure]

  [rules composer]

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Application Examples} \label{sec:evaluation}

[smartlic]

[media follow me]

[prenda]

@@

>>>>

% *** SPECIALIZED LIST PACKAGES ***
%
\usepackage{algorithmic}
% \renewcommand{\algorithmicrequire}{\textbf{Input:}}
\newcommand{\algorithmicforeach}{\textbf{for each}}
\newcommand{\algorithmiconevent}{\textbf{on event received}}
\newcommand{\algorithmicendevent}{\algorithmicend\ \algorithmionevent}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

@@

% With the purpose of evaluating the proposal we implemented the concepts 
% and enable the context interpretation and 
%  even without the need of purchasing all those devices
%  (with a touch screen interface)
% The abstractions provided, and the management services conceived, allow developers to handle resources spread within the intelligent environment in a simple and homogeneous way, and enhances the capability of implementing context sensible applications, while still keep the flexibility on developing such applications. 
% Above the platform we implemented a management and prototyping interface that uses a tablet device to allow the testing of applications running on different configurations of the environment, which may be loaded with simulators or the representation of the resources. The interface also allows the composition of context rules either by developers or non trained users.

@@

% [architecture picture]
\begin{figure}[ht]
\centering
\includegraphics[width=3.5in]{figures/architecture-refactored-v2.pdf}
\caption{Framework Layer Architecture}
\label{fig:arch}
\end{figure}

@@

    %[talk about stub intantiation calls]
    %TODO: example of CV and OP? Maybe it is worth add some on the text
    To perform any interaction with a RA that is not locally instantiated, the UbiApp must use a stub (also called proxy) that is bound to the referred RA. This stub implements the same interface as the RA, i.e. the same CVs and OPs, which are methods, from a Object-Oriented Programming perspective. With the stub on hands, the application can invoke remote methods, as if they were locally invoked. These remote methods may be to retrieve a CV or to trigger an OP. The Figure~\ref{fig:rpc} represents this direct, or synchronous, communication. The steps to achieve a direct communication, as represented by the figure, are as follows:
    \begin{enumerate}
     \item The UbiApp calls a method from the stub 
     \item The call is forwarded to the RA instance through the network in the second step
     \item This instance can interact with the device (the lamp, in case)
     \item The result or the acknowledge is then returned to the caller
    \end{enumerate}

%     In the first step the UbiApp calls a method from the stub whose call is forwarded to the RA instance throught the network in the sencond step, this instance can interact with the device (the lamp, in case) and return an answer or an aknowledge. In this example, the communication could be used to retrieve come CV status from the lamp such as ``is on'' or ``light intensity''

    \begin{figure}[ht]
    \centering
    \includegraphics[width=1.8in]{figures/comm-rpc.pdf}
    \caption{Direct communication} 
    \label{fig:rpc}
    \end{figure}

    It is possible to acquire the status of a CV directly, however it can also be done indirectly via the pub-sub mechanism, what makes it easier for developers being aware of CV change right away. This indirect, or asynchronous, communication is performed in two stages: the subscription (Figure~\ref{fig:pubsub}.a) and the notification (Figure~\ref{fig:pubsub}.b). During the first stage, represented by the Figure~\ref{fig:pubsub}.a, we have the following steps:
    \begin{enumerate}
     \item The UbiApp sets some RA (``RA X'') to subscribe to another, referred by its RANS (``Lamp RA'') and in a specific CV (``Is on'')
     \item Then the first RA requires to be subscribed to the second, through the network
     \item The second RA accepts the subscription and is bound to the device
    \end{enumerate}

    The second stage happens when the first RA is already subscribed to the second and a notification routine is triggered, what was represented by the Figure~\ref{fig:pubsub}.b and works as follows:
    \begin{enumerate}
     \item The device has its context changed, what may be caused directly (e.g., a user turning the lamp on) or through the RA (i.e. an application, another RA, or rule, calling some RA's OP which caused the change of context)
     \item Then the RA that represents the device notifies all subscriptors to that CV (e.g., ``Is on'' from this Lamp RA)
     \item The subscriptor may call some callback method of the UbiApp
    \end{enumerate}

    \begin{figure}[ht]
    \centering
    \includegraphics[width=3.4in]{figures/comm-pubsub.pdf}
    \caption{Indirect communication} 
    \label{fig:pubsub}
    \end{figure}

@@

%[David]
The conceptual framework proposed at this work was tested through a reference implementation called \mbox{SmartAndroid} Project\footnote{For more information visit www.tempo.uff.br/smartandroid}. The project includes a programming API to help the developers in constructing UbiApps as well as offers basic management services to provide resource registration, discovery and localization. The SmartAndroid has been developed over Google Android platform as we believe that in a near future it will be widely used in sensors and devices embedded in furnitures and home appliances.
% [we have implemented most of the framework features, thought much more remain to be done]

%proposta de uso de marketplace
% A exposição de dados por meio de \textit{Content Provider} permite a integração de aplicativos externos ao servidor. Estes aplicativos podem ser dispostos em um \textit{marketplace} (http://developer.android.com/distribute/open.html) permitindo que tanto usuários comuns quanto programadores incrementem o ambiente com novas aplicações ubíquas. 
The SmartAndroid middleware contemplates the possility of running multiple applications in the same IE. Thus, focusing on the user and inspired by the Google Play -- which is a application market for Android devices --, we have as a future work the creation of a SmartAndroid Market~\cite{google2013}, where users will be able to easily download and start different applications. Also similar to the Google Play, a Manifest file will require permission for accessing the domains, what will allow users to maintain control over what off-the-shelf applications have permission to know and to do.
% TODO: talk about market place
[security concern on store: helps the discovery of malware before the user download]

@@

  [initialization]
%   O SmartAndroid é um aplicativo que inicializa o SGAR, repositório e serviços de comunicação. Este aplicativo deve ser executado em uma única unidade do ambiente, deixando o Deamon de Comunicação em estado de espera até receber requisições de outras aplicações. Em outras unidades (clientes) constam os aplicativos instalados com a biblioteca do SmartAndroid, a qual é necessária para executar o serviço de comunicação e \textit{proxys} dos serviços do SGAR. Assim cada unidade pode consultar sobre IARs, através do SDR, obtendo seus ODARs, cada um com o respectivo RANS, o qual é utilizado pelo \textit{proxy} como referência de acesso.

  with the RANS 
  RPC search

  [annotation]

  \subsection{Context Rules} \label{subsec:rules}
%   [implementation of context rules]
  The SmartAndroid supports the context interpretation by providing mechanisms to create context rules. The context rules life cycle has two major phases: composition and running. The composition phase corresponds to the instantiation of a CI and of an actuator and the binding of both through subscription. The running phase is when the CI and the actuator(s) are already instantiated as the other RAs that feed the IE with context information, and they are able to exchange messages and perform their job.
  %   Support the interpretation of the context framework enables a separation of concerns, where developers abstract implementation of contextual rules and focus on the application logic.

  At the implementation level, a context rule can be described in a JSON structured file (or byte stream) following a specific format. In the first part an expression must be defined setting the logical operat
  To separate the context rules creation from the

%   O IC é encapsulado em um AR, estendendo, assim, suas funcionalidades. Sua criação ocorre quando é definida uma expressão lógica no objeto do IC recém instanciado, segundo um padrão próprio adotado em arquivos tipo JSON, e um \textit{parser} constrói uma estrutura lógica interna, como mostrado na Figura~\ref{fig:inpt_arq}. A estrutura lógica abrigada pelo IC, contém uma estrutura em árvore que representa a expressão e contém não só a lógica entre as VCs das condições, como também os valores atualizados das mesmas. Enquanto em funcionamento, o IC recebe atualizações das VCs, tratando-as em seguida para atualizar sua estrutura lógica. O módulo de avaliação é notificado de mudanças da estrutura lógica e avalia se a expressão se tornou verdadeira ou falsa. Se a expressão alcançou um estado verdadeiro, considerando as possíveis temporizações, notifica os atuadores.

@@

  \subsection{Interface for Management and Prototyping} \label{subsec:ipgap}
  % interface de prototipagem e gerenciamento de aplicações pervasivas
  % prototyping and management of pervasive applications interface
  % pervasive applications' management and prototyping interface
  % PAMPI
  % management and prototyping interface MPI
  [description]\\

  %[David]
  The main purpose of the [IPGAP] is provide to developers a hybrid environment where simulated and real resources can interact to each other transparently, enabling their users to construct UbiApps as well as test and evaluate them through a wide variety of scenarios. Thus RAs from sensors, actuators and household appliances can be graphically represented at our tool, along with a schematic map from the intelligent environment so that some of the resources aspects as its context and location can be easily visualized on the map, helping in application debugging.

  Besides, [IPGAP] operate with two different interfaces: the \textit{developer interface}, already mentioned, and the \textit{end-user interface}. In the end-user interface, users in general, i.e. non-technical users, are able to use [IPGAP] to manage their intelligent environment remotely using a mobile device, such as a tablet. Through an intuitive GUI, end-users can not only visualize the state of their home devices but control them. For instance, it's possible turning off a stove burner, change the air-conditioner temperature, and so on. By touching the device icon in the screen, an interface that presents the device's main functions is shown, where the user can operate it as the tool were a remote control.

  % [main figure]
  \begin{figure}[!t]
  \centering
  \includegraphics[width=3.5in]{figures/map-forniture-cv-menu-big.pdf}
  \caption{Management and Prototyping Interface} %[IPGAP]
  \label{fig:ipgap}
  \end{figure}
  %FIXME: convert portuguese text to english

%   [rules composer]
  %TODO Think about a better rule example!!
  In addition, end-users can compose context rules in a hi-level way by graphically combining the CVs from any RA and logical connectives, forming logical expressions as \textit{``if the TV from living room \textbf{is on}, \textbf{and} nobody has been present at the living room for 15 minutes, \textbf{then} turn off that TV''}. To compose this rule the user have to, first of all, select the rule composer mode in the [IPGAP]. Then, he must touch the TV icon at the living room representation on the map, and choose the desired CV -- in that case, the CV that refers if the TV is turned on or not. After that, a dialog box showing all possible logical operators is shown, so that the user can select one of them as the \textbf{AND} operator. The user can repeat this process, adding others CVs and operators until the expression is complete. Finally, the rule can be associated to one or more actions, provided by actuators.
  %FIXME: here ``AND'' is bold, but before it was just between quotation marks

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Example Applications} \label{sec:evaluation}
Some applications have been built to explore the potential of the framework for building UbiApps. Although the example applications are simple, and alone does not justify the deploy of the SmartAndroid in the environment, they have allowed us to validate the features provided by the framework and achieved by the SmartAndroid.

% Among these applications we will

% the SmartLiC, the MediaFollowMe, remote control type applications where resources are controlled remotely by another resource (in this case, a smartphone ), among other applications. A future work is to install a system in Ambi home care health, as in ~ \ cite {} carvalho2010sciads.

  \subsection{Tic-Tac-Toe}
  After the implementations of the basic features of the framework, including the communication mechanisms, we have implemented a distributed Tic-Tac-Too game, where a player can defy another player (or AI player) placed anywhere in the IE. We have used as devices two smartphones.

  In this implementation we have created RAs for both the players and the board. The players subscribe to the board and vice-versa, thus each movement performed by the player is set on the board through RPC, commands from this player are locked and the movement is communicated to the other player, the same happens to the second player, and so on. One of the goals of this implementation was to verify the capability of the framework to make the construction of distributed systems easier.

  \subsection{SmartLiC}
  % [smartlic]
  The Smart Light Control (SmartLic) application aims at the clever use of lighting in homes in order to save power consumption, since it is responsible for the largest range of consumption in residential environments. The application uses presence sensors to identify the location of people inside each room, and decide which lamps should be turned on and off. Hence it prevents the forgetting of lamps turned on, and may lit some lamp if a user enters a room at night and if nobody is at sleep in that room. 
  % ref to the 'largest range of consumption in residential environments'?

  The [IPGAP] has allowed us to test this applications. We instantiated lamps simulators for each room and an avatar for a user, and also a simulator for day light sensor. Next context rules were created by filling a skeleton of a rule written in json, what caused the instantiation of a CI and actuator for each room.  If the avatar is dragged, in the [IPGAP], from a room and dropped in another, it ma
%   O Código 2 ilustra uma possível implementação da interpretação de contexto pela aplicação \mbox{SmartLiC}. Na etapa de inicialização, a aplicação utiliza o SLR para localizar os cômodos (Linha 2), e obter a lâmpada e o sensor de presença (``lampAR'' e ``presAR'') de cada um, utilizando a busca do SDR (Linhas 4 e 5). Para evitar a repetição do código da regra, suas referências e a respectiva ação a ser desempenhada (desligar a lâmpada referenciada -- ``lampAR.turnOff'') são adicionadas na lista de configurações (``configList'') (Linha 6). A regra é criada em ``rule'' como: ``\mbox{\emph{lâmpada ligada E o cômodo desocupado POR T$_{max}$ unidades de tempo}}'', e desempenha uma relação lógica entre as diferentes VCs e a temporização da regra (Linha 7). Em seguida, subscreve-se às VCs dos ARs de interesse, começando o processamento da regra (Linha 8).

%   Na etapa de avaliação, também descrita no Código 2, é definido um bloco de comando para o tratamento das notificações recebidas dos ARs de interesse (lâmpadas e sensores de presença subscritos). Uma vez recebido um evento com a atualização de uma VC de um dos ARs (``ar\_ref'') (Linha 11), é selecionada a configuração de ``configList'' respectiva (Linha 12). Segue-se a avaliação da regra (\textit{``evaluate}''), utilizando a configuração obtida, e considerando os temporizadores (Linha 13). A etapa de avaliação é bloqueante no caso da existência de temporizadores associados, e a invocação das ações (OPs) ocorre na mesma \textit{thread} que a avaliação. A temporização ocorre enquanto a expressão da regra é válida, se eventos chegarem antes do fim da temporização e isto fizer com que a expressão da regra seja invalidada, então o temporizador é finalizado e as ações não são disparadas.

%   \begin{lstlisting}[caption=Context Interpretation by the SmartLiC, mathescape] 
%   $\textit{//Initialization}$
%   placeList = SLR.getPlaces() $\textit{//Get rooms}$
%   for each place from placeList: $\textit{//Selects a room from the list}$
%       lampAR = SDR.search(Lamp, place) $\textit{//Get lamps from the room}$
%       presAR = SDR.search(Presence, place) $\textit{//Get presence sensor from the room}$
%       configList.add(lampAR, presAR, lampAR.turnOff) $\textit{//Stores configuration}$
%   rule = Lamp.on $\land$ $\lnot$(Presence.occupied) $\land$ T > T$_{max}$ $\textit{//Rule function}$
%   subscribeTo(configList) $\textit{//Subscribes to lamps and presence sensors}$
% 
%   $\textit{//Context rule evaluation}$
%   on event received (ar_ref, vc, value): $\textit{//Notification received}$
%       config = configList(ar_ref, vc) $\textit{//RA configuration selection}$
%       if (evaluate(rule, config)) then: $\textit{//Evaluates rule with configuration}$
% 	  invoke(config.getActions()) $\textit{//Calls action}$
%   \end{lstlisting}

%   \begin{algorithmic}
%   \STATE $asdf$
%   \FORALL{$asdf$ in $asdf$}
%     \STATE $asdf$
%   \ENDFOR
%   
%   \ONEVENT{$(ra, cv, value)$}
%     \STATE $asdf$
%   \ENDEVENT
%   \end{algorithmic}

%   \begin{algorithmic}
%   \STATE $rule \leftarrow Lamp.on \land \lnot(Presence.occupied) \land T > T_{max}$
%   \STATE $placeList \leftarrow RLS.getPlaces()$
%   \FORALL{$place$ in $placeList$}
%     \STATE $lampRA \leftarrow RDS.search(Lamp, place)$
%     \STATE $presRA \leftarrow RDS.search(Presence, place)$
%     \STATE $rule$
%     \STATE $RAs.add(presRA)$
%     \STATE $new ContextInterpreter()$
%   \ENDFOR
%   \STATE $rule \leftarrow Lamp.on \land \lnot(Presence.occupied) \land T > T_{max}$
%   \STATE $subscribeTo(RAs)$
%   \end{algorithmic}
%   
%   
%   \begin{algorithmic}
%   \STATE $placeList \leftarrow RLS.getPlaces()$
%   \FORALL{$place$ in $placeList$}
%     \STATE $lampRA \leftarrow RDS.search(Lamp, place)$
%     \STATE $presRA \leftarrow RDS.search(Presence, place)$
%     \STATE $configList.add(lampRA, presRA, lampRA.turnOff)$
%   \ENDFOR
%   \STATE $rule \leftarrow Lamp.on \land \lnot(Presence.occupied) \land T > T_{max}$
%   \STATE $subscribeTo(configList)$
%   \end{algorithmic}

  Another feature that we will implement is a house security mode, that users may activate if they are travelling or just have left the house alone. The basic operation is to turn lamps on and off as if users were at home, thus deceiving potential thieves. The choices of which lamps to turn on and off will follow the pattern contained in a log file, to mimic a realistic scenario.

  \subsection{Remote Controls}
  Considering that the users would want to have some control over the resources within the IE, we have implemented the remote controls to interact with them specifically.

  \begin{figure}[!t]
  \centering
  \includegraphics[width=3.5in]{figures/beagle-lamp.pdf}
  \caption{Remote Control} %[IPGAP]
  \label{fig:remotecontrol}
  \end{figure}

  \subsection{Prenda Game}
%   um agente pode saber quais recursos mais próximo de um ponto de referência específico
  The Prenda is a ubiquitous game implemented for the SmartAndroid where the goal is to find a gift hidden somewhere in the house. The player moves inside the house and gets tips on where to find the gift while competing with others, the one who finds the gift first is the winner. This game has a different concept of other common digital games, since it explores elements in the IE and thus offers a different experience for kids.

  This game was implemented with simulated resources and using the [IPGAP]. The players were represented with avatars and their location were updated with the drag and drop of the avatars over the map.

  The player has a layout map of the house where you can find the move to attach the extent that the movement occurs, the game informs the person's subjective closeness with the gift (by the number of points in the message text) to that is found. The game allows the participation of multiple players for the environment, to get a winner the gift is repositioned to the new game. The game takes place in a simulated environment, using IPGAP to simulate the movement of avatars of people participating.

  \subsection{MediaFollowMe}

  The \mbox{\textbf{MediaFollowMe}} application enables the user to be followed by any kind of media stream (e.g. video, audio, images) while he/she walks inside the intelligent environment. By the time the user launch the \mbox{MediaFollowMe} app and choose the device which the media stream he wants to be followed, the app will take care of the user's location changes and automatically play, pause or migrate the media to the closest device capable of displaying it.

  To illustrate that, consider a user watching a movie on TV through a blu-ray player in his bedroom. When the user leaves his bedroom, the system identifies this action by an RFID tag and fires an event, as well as when he arrives in any room. As the app is interested in these events, it receives a notification informing that the user left/arrived some room (and which room) and can migrate the media to the proper device, preferable inside that room.

@@

% REMAIN TODO TASKS
% write about Gaia, Gator Tech and Context Toolkit
% remove SoA references
% security concerns on marketstore
% evaluate examples spread in text
% araujo is a paper in PT... find another one better
% fix figure that contains PT text
% finish related work
% finish example applications
